---
title: "DATA607_Week3"
author: "Logan Thomson"
date: "February 13, 2016"
output: html_document
---

###Setting Up Database Connection
After setting up the tables in SQL, we can use the `RMySQL` package to connect to the MySQL database and use an SQL script just as we would in MySQL.  

```{r}
library('RMySQL')
drv <- dbDriver('MySQL')
con <- dbConnect(drv, user="root", password="", dbname = "tb",
                 host = "localhost")
```

Using the combination of the MySQL driver stored in `drv` and the database connection, stored in `con`, we can view the tables in our database
```{r}
dbListTables(con)
```

As well as view the fields in a particular table
```{r}
dbListFields(con, "tb")
```

#### Querying the Database  
Following is the code that does most of the work. This is the same `SELECT` statement that wold retrieve the data from each table and combine it together to return the data. The SQL script is stored in the `query` object.

```{r, warning=FALSE}
query <- "SELECT t.country, t.year, sum(t.cases) AS cases, p.population,
(t.cases / p.population) * 1000 AS rate
FROM tb_cases AS t
INNER JOIN population AS p
ON p.country = t.country
AND p.year = t.year
WHERE cases is not null
GROUP BY t.country, t.year"

cases_per_1000 <- dbGetQuery(con, query)
```

Using `dbGetQuery` and calling our `con` connection object and the SQL script stored in `query`, we can store the retrieved data in an object, which I have called `cases_per_1000`.  Because the rate returned by dividing cases by population was small, I used a multiple to return a more managable number.
```{r, warning=FALSE}
dbDisconnect(con)
dbUnloadDriver(drv)
```

###Creating Data Frame  
To manipulate the data in R, the returned SQL data can be stored in a data frame, which I have called `cases_df`.  Below is a preview of the first few rows:  
```{r}
cases_df <- data.frame(cases_per_1000)
head(cases_df)
```

And  here are the summary statistics for the data:  

```{r}
summary(cases_df)
```

###Plotting Data  

Lets make a quick histogram of the rate data to see what we are working with:
```{r, fig.width=10, fig.height=6}
hist(cases_df$rate, breaks=30)
```

In most years, countries have less than 1 case per 1000, and in more severe cases, there are several cases per 1000 people, which is a relatively high rate.  Still, for the rates below 1 per 1000, there is a lot of variance.  

Let's take a look at two countries that greatly differ in their overall economies, the United States, and Angola (TB rates in Africa are generally higher than other parts of the world):  

```{r, fig.width=10}
library(ggplot2)
US_vs_Angola <- subset(cases_df, country == 'Angola'| country == 'United States of America')
ggplot(US_vs_Angola, aes(x=US_vs_Angola$year, y=US_vs_Angola$rate)) + geom_point(aes(color = US_vs_Angola$country)) + labs(title='TB Cases United States vs. Angola', x="Year", y="TB Rate per 1000 People")
```
Though both have rates that are below 1 case per 1000 people, the rate of TB cases in Angola is much higher than that of the rate in the United States, which is virtually zero.
